---
Title: 'Class Weekend Project'
---
**Livingston Akoma
**DAB402 - 0773239
---

#### 1. Read the csv file (USDA_Clean.csv) in the folder and assign it to a data frame. (3 points)

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)


class_file <- read_csv("/Users/livingstonakoma/OneDrive/St. Claire College/Semester 4/Capstone Class_ClassFiles_09_08_21/USDA_Clean.csv")

view(class_file)
```

#### 2. Check the datatypes of the attributes. (3 points)

```{r}
sapply(class_file, class)
```

#### 3. Visualize the correlation among Calories, Protein, Total Fat, Carbohydrate, Sodium and Cholesterol. (7 points)

```{r}

my_data <- class_file[,c("Calories","Protein","TotalFat","Carbohydrate","Sodium","Cholesterol")] 
correlation <- cor(my_data) 

round(correlation,2) 

```

#### 4. Is the correlation between Calories and Total Fat statistically significant? Why? (7 points)

```{r}

cor(my_data$Calories,my_data$TotalFat) 

#correlation between Calories and Total Fat is statistically significant
#because value of correlation is close to 1(it is strong a positive correlated)
```

#### 5. Create a Linear Regression Model, using Calories as the dependent variable Protein, Total Fat, Carbohydrate, Sodium and Cholesterol as the independent variables. (7 points)

```{r}
mod <- lm(Calories ~ Protein + TotalFat + Carbohydrate + Sodium + Cholesterol , data = my_data  ) 

summary(mod) 
```

#### 6. Write the Linear Regression Equation, using Calories as the dependent variable whereas Protein, TotalFat, Carbohydrate, Sodium and Cholesterol as the independent variables. (7 points)

```{r}

#  Calories = coef(mod)[1] + coef(mod)[2]*Protein + coef(mod)[3]*TotalFat   +coef(mod)[4]*Carbohydrate + coef(mod)[5]*Sodium + coef(mod)[6]*Cholesterol 

#  Calories = 3.9882753 + 3.9891994*Protein + 8.7716980*TotalFat +3.7432001*Carbohydrate + 0.0003383*Sodium + 0.0110138*Cholesterol
```

#### 7. Which independent variable is the least significant? Why? (7 points)

```{r}

# Sodium is the least significant because its p-value (0.122 ) > 0.05. 

```

#### 8. A new product is just produced with the following data:                                                            Protein=0.1, TotalFat=35, Carbohydrate=405, Sodium=440, Cholesterol=70, Sugar=NA, Calcium=35, Iron=NA, Potassium=35, VitaminC=10,	VitaminE=NA, VitaminD=NA.                                                                                                         Based on the model you created, what is the predicted value for Calories? (7 points)

```{r}

Calories = coef(mod)[1] + coef(mod)[2] * 0.1 + coef(mod)[3] * 35 +coef(mod)[4] * 405+ coef(mod)[5] * 440 + coef(mod)[6] * 70 

Calories

```

#### 9. If the Sodium amount increases from 440 to 44440 (10000% increase), how much change will occur on Calories in percent? Explain why? (7 points)

```{r}


pred_Calories = coef(mod)[1] + coef(mod)[2] * 0.1 + coef(mod)[3] * 35 +coef(mod)[4] * 405+ coef(mod)[5] * 440 + coef(mod)[6] * 70 
predNew_Calories = coef(mod)[1] + coef(mod)[2] * 0.1 + coef(mod)[3] * 35 +coef(mod)[4] * 405 + coef(mod)[5] * 44440 + coef(mod)[6] * 70 

percentChange_Calories = ((predNew_Calories - pred_Calories )/ pred_Calories) * 100 

percentChange_Calories 


# If we increase sodium to 1 unit it will increase calorie by 0.03383% and rest of the independent variables are constant. Slope of sodium is positive the change is positive.

```


#### 10. A study of primary education asked elementaty school students to retell two book articles that they read earlier in the week. The first (Article 1) had no pictures, and the second (Article 2) was illustrated with pictures. An expert listened to recordings of the students retelling each article and assigned a score for certain uses of language. Higher scores are better. Here are the data for five readers in this study:

#### Article 1   0.40   0.72   0.00   0.36   0.55

#### Article 2   0.77   0.49   0.66   0.28   0.38


#### A) What are $H_0$ and $H_a$ ? (5 points)

```{r}

# HO : mean(Article 1 )  =  mean(Article 2) 

# Ha : mean(Article 1 ) <  mean(Article 2) 

```

#### B) Is this a paired or unpaired experiment? (5 points)

```{r}
# it is a paired experiment. 
```

#### C) Based on your previous answer, which nonparametric test statistic would you use to compare the medians of Article 1 and Article 2. (5 points) 

```{r}

# wilcoxon signed rank test 

```

#### D) Use a nonparametric test statistic to check if there is a statistically significant difference between the medians of Article 1 and Article 2. (5 points) 

```{r}

Article_1 <- c(0.40, 0.72, 0.00, 0.36, 0.55) 
Article_2 <- c(0.77, 0.49, 0.66, 0.28, 0.38)  

wilcox.test(Article_1,Article_2 , paired = T ,alternative = "less") 

##  Wilcoxon signed rank test 

## data:  Article_1 and Article_2 

## V = 6, p-value = 0.4062 

## alternative hypothesis: true location shift is less than 0 

```

#### E) Will you accept or reject your Null Hypothesis? ($\alpha = 0.05$) Do illustrations improve how the students retell an article or not? Why? (5 points)

```{r}

# So p-value is greater than 0.05.Thus we do not have enough evidence to reject null hypothesis 

#students are performing the same with or without illustration and illustrations do not improve how the students retell an article. 

```

#### 11. Two companies selling toothpastes with the lable of 100 grams per tube on the package. We randomly bought eight toothpastes from each company A and B from random stores. Afterwards, we scaled them using high precision scale. Our measurements are recorded as follows:

#### Company A: 97.1   101.3   107.8   101.9   97.4    104.5   99.5    95.1
#### Company B: 103.5  105.3   106.5   107.9   102.1   105.6   109.8   97.2


#### A) Is this a paired or unpaired experiment? (5 points)

```{r}

Company_A<- c(97.1,101.3,107.8,101.9,97.4,104.5,99.5,95.1) 
Company_B <- c(103.5,105.3,106.5,107.9,102.1,105.6,109.8,97.2) 
 
# it is a unpaired experiment 

```

#### B) Based on your previous answer, which nonparametric test statistic would you use to compare the medians of Company A and Company B. (5 points) 

```{r}


# Wilcoxon's Rank Sum Test 


```

#### C) Use a nonparametric test statistic to check if there is a statistically significant difference between the medians of Company A and Company B. (5 points) 

```{r}


wilcox.test(Company_A,Company_B) 


##  Wilcoxon rank sum test 
## data:  Company_A and Company_B 
## W = 13, p-value = 0.04988 
## alternative hypothesis: true location shift is not equal to 0 

# for 95% confident: p-value(0.04988) < 0.05. so we reject the null hypothesis and we accept that, there is significant difference  difference between packaging of these two companies. 

```

#### D) Will you accept or reject your Null Hypothesis? ($\alpha = 0.05$) Are packaging process similar or different based on weight measurements? Why? (5 points)

```{r}


# for 95% confident: p value < 0.05. so we reject the null hypothesis and we accept that, there is significant difference  difference between packaging of these two companies. 


```
